---
title: "A Nonparametric Bayesian Shrinkage Estimator for Binomial Data under Informative Sample Sizes"
author: "Lufeiya Liu (lul115@pitt.edu)"
date: "Spring 2026"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: readable
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6, cache = TRUE)
set.seed(42)
```

# Overview

This project reproduces and extends **Zhang & Liu (2012)**, "Nonparametric hierarchical
Bayes analysis of binomial data via Bernstein polynomial priors" (*Canadian Journal of
Statistics*). The standard Bernstein-Dirichlet Process (BDP) treats `N_i` as fixed and
exchangeable. We add a log-normal observation process

$$\log(N_i) \mid \theta_i \sim \mathrm{Normal}(\alpha + \beta\theta_i,\, \sigma^2)$$

to jointly model the dependence between sample size and latent success probability.
When `beta = 0`, the model reduces exactly to Zhang and Liu's standard BDP.

**Headline results.**

- Baseball (Brown's 2005 MLB): TSE* = 0.455 at `K_max = 1200` (27% improvement over standard BDP, 10% over NPEB).
- Stack Overflow (n = 2,491): TSE* = 0.766 at `K_max = 200`, lowest among methods evaluated.
- Simulation (5 scenarios, 30 replicates): crossover at `beta ≈ 1`; informative BDP dominates above the threshold.

# Section 1: Core Functions

## 1.1 Standard BDP MCMC (Zhang & Liu 2012, Procedure I)

Implements the Gibbs sampler from Zhang & Liu (2012, Section 4.1). State space is
`(Z_1, ..., Z_n, k, M)`. Returns Rao-Blackwellized point estimates
`theta_hat_i = E[(j_i + Y_i) / (k + 1 + N_i) | data]`.

```{r bdp-mcmc}
bdp_mcmc <- function(y, N, n_iter = 20000, n_burn = 5000, K_max = 200,
                     M_grid = seq(0.1, 10, by = 0.1), k_init = 30, seed = 42) {
  set.seed(seed)
  n <- length(y)
  stopifnot(length(N) == n, n > 0)

  # Initialization (data-informed to avoid k=1 trap)
  k <- k_init
  M <- 2.0
  theta_mle_init <- (y + 0.5) / (N + 1.0)
  Z <- pmin(pmax(theta_mle_init + rnorm(n, 0, 0.02), 0.001), 0.999)
  j_lab <- pmin(ceiling(Z * k), k)
  j_lab[j_lab < 1] <- 1

  # Storage
  n_total <- n_iter + n_burn
  k_trace <- integer(n_total)
  M_trace <- numeric(n_total)
  theta_grid <- seq(0.005, 0.995, length.out = 300)
  density_sum <- numeric(length(theta_grid))
  theta_hat_sum <- numeric(n)
  n_kept <- 0

  cat(sprintf("Standard BDP: %d iterations (%d burn-in), K_max=%d, n=%d\n",
              n_iter, n_burn, K_max, n))

  for (iter in 1:n_total) {

    # Step I.a: Sample k
    log_pk <- numeric(K_max)
    for (kk in 1:K_max) {
      jj <- pmin(pmax(ceiling(Z * kk), 1), kk)
      log_pk[kk] <- sum(lbeta(jj + y, kk + 1 - jj + N - y) -
                        lbeta(jj, kk + 1 - jj))
    }
    log_pk <- log_pk - max(log_pk)
    pk <- exp(log_pk); pk <- pk / sum(pk)
    k <- sample.int(K_max, 1, prob = pk)
    j_lab <- pmin(pmax(ceiling(Z * k), 1), k)

    # Step I.b: Sample Z_i (Polya urn)
    alpha_jk <- 1 / k
    jvec <- 1:k
    log_lik_mat <- outer(jvec, seq_len(n), function(j, i) {
      lbeta(j + y[i], k + 1 - j + N[i] - y[i]) - lbeta(j, k + 1 - j)
    })
    nj_full <- tabulate(j_lab, nbins = k)

    for (i in 1:n) {
      old_j <- j_lab[i]
      nj_i <- nj_full
      nj_i[old_j] <- nj_i[old_j] - 1

      log_pj <- log(M * alpha_jk + nj_i) + log_lik_mat[, i]
      log_pj <- log_pj - max(log_pj)
      pj <- exp(log_pj); pj <- pj / sum(pj)

      j_new <- sample.int(k, 1, prob = pj)
      nj_full[old_j] <- nj_full[old_j] - 1
      nj_full[j_new] <- nj_full[j_new] + 1
      j_lab[i] <- j_new
      Z[i] <- runif(1, (j_new - 1) / k, j_new / k)
    }

    # Step I.c: Sample M
    r <- length(unique(j_lab))
    log_pM <- sapply(M_grid, function(Mv) {
      r * log(Mv) - sum(log(Mv + 0:(n - 1)))
    })
    log_pM <- log_pM - max(log_pM)
    pM <- exp(log_pM); pM <- pM / sum(pM)
    M <- M_grid[sample.int(length(M_grid), 1, prob = pM)]

    k_trace[iter] <- k
    M_trace[iter] <- M

    # Accumulate post-burn-in (Rao-Blackwellized)
    if (iter > n_burn) {
      n_kept <- n_kept + 1
      nj <- tabulate(j_lab, nbins = k)
      W_rb <- (M * alpha_jk + nj) / (M + n)
      for (g in seq_along(theta_grid)) {
        density_sum[g] <- density_sum[g] +
          sum(W_rb * dbeta(theta_grid[g], 1:k, k + 1 - (1:k)))
      }
      theta_hat_sum <- theta_hat_sum + (j_lab + y) / (k + 1 + N)
    }

    if (iter %% 5000 == 0)
      cat(sprintf("  iter %d/%d  k=%d  M=%.1f  r=%d\n",
                  iter, n_total, k, M, r))
  }

  list(
    theta_grid  = theta_grid,
    density_est = density_sum / n_kept,
    theta_hat   = theta_hat_sum / n_kept,
    k_samples   = k_trace[(n_burn + 1):n_total],
    M_samples   = M_trace[(n_burn + 1):n_total]
  )
}
```

## 1.2 Informative BDP MCMC (Log-Normal Observation Process)

Extends the standard BDP with a log-normal observation process for `N_i` given
`theta_i`. The full state is `(theta, Z, k, M, alpha, beta, sigma^2)`. Updates:

- `theta_i` via Metropolis-Hastings (Beta proposal absorbing binomial + BDP factors;
  the MH ratio corrects only for the log-normal factor)
- `(Z_i, k, M)` via Gibbs (same as standard BDP)
- `(alpha, beta)` via conjugate bivariate normal
- `sigma^2` via conjugate inverse-gamma

When `beta = 0`, the log-normal factor is constant in `theta_i`, every MH proposal
is accepted, and the sampler reduces to Zhang and Liu's Procedure I.

```{r informative-bdp-lognormal}
informative_bdp_lognormal <- function(y, N, n_iter = 20000, n_burn = 5000,
                                       K_max = 200,
                                       M_grid = seq(0.1, 10, by = 0.1),
                                       alpha_prior_mean = 4, alpha_prior_var = 100,
                                       beta_prior_mean = 0,  beta_prior_var = 100,
                                       sigma2_prior_a = 1,   sigma2_prior_b = 1,
                                       seed = 42) {
  set.seed(seed)
  nn <- length(y)
  stopifnot(length(N) == nn, nn > 0)

  logN <- log(N)

  # Initialization
  k <- 30
  M <- 2.0
  theta <- pmin(pmax((y + 0.5) / (N + 1.0), 0.01), 0.99)
  Z <- pmin(pmax(theta + rnorm(nn, 0, 0.02), 0.001), 0.999)
  j_lab <- pmin(pmax(ceiling(Z * k), 1), k)

  # OLS init for observation-process parameters
  fit0 <- lm(logN ~ theta)
  alpha_op <- as.numeric(coef(fit0)[1])
  beta_op  <- as.numeric(coef(fit0)[2])
  sigma2   <- summary(fit0)$sigma^2
  cat(sprintf("  OLS init: alpha=%.2f, beta=%.2f, sigma2=%.2f\n",
              alpha_op, beta_op, sigma2))

  theta_accept <- 0
  theta_total  <- 0

  # --- Storage ---
  n_total <- n_iter + n_burn
  k_trace      <- integer(n_total)
  M_trace      <- numeric(n_total)
  alpha_trace  <- numeric(n_total)
  beta_trace   <- numeric(n_total)
  sigma2_trace <- numeric(n_total)
  theta_grid   <- seq(0.005, 0.995, length.out = 300)
  density_sum  <- numeric(length(theta_grid))
  theta_hat_sum <- numeric(nn)
  n_kept <- 0

  cat(sprintf("Log-Normal Informative BDP: %d iter (%d burn-in), K_max=%d, n=%d\n",
              n_iter, n_burn, K_max, nn))

  for (iter in 1:n_total) {

    # Step 1: theta_i via MH (Beta proposal, log-normal correction)
    for (i in 1:nn) {
      j <- j_lab[i]
      a_post <- j + y[i]
      b_post <- k + 1 - j + N[i] - y[i]
      theta_star <- rbeta(1, a_post, b_post)
      resid_star <- logN[i] - alpha_op - beta_op * theta_star
      resid_curr <- logN[i] - alpha_op - beta_op * theta[i]
      log_ratio  <- (-resid_star^2 + resid_curr^2) / (2 * sigma2)
      if (log(runif(1)) < log_ratio) {
        theta[i] <- theta_star
        theta_accept <- theta_accept + 1
      }
      theta_total <- theta_total + 1
    }
    Z <- pmin(pmax(theta + rnorm(nn, 0, 0.001), 0.001), 0.999)
    j_lab <- pmin(pmax(ceiling(Z * k), 1), k)

    # Step 2: j_i (Polya urn)
    alpha_jk <- 1 / k
    nj_full <- tabulate(j_lab, nbins = k)
    for (i in 1:nn) {
      old_j <- j_lab[i]
      nj_i <- nj_full; nj_i[old_j] <- nj_i[old_j] - 1
      log_pj <- log(M * alpha_jk + nj_i) +
                dbeta(theta[i], 1:k, k + 1 - (1:k), log = TRUE)
      log_pj <- log_pj - max(log_pj)
      pj <- exp(log_pj); pj <- pj / sum(pj)
      j_new <- sample.int(k, 1, prob = pj)
      nj_full[old_j] <- nj_full[old_j] - 1
      nj_full[j_new] <- nj_full[j_new] + 1
      j_lab[i] <- j_new
      Z[i] <- runif(1, (j_new - 1) / k, j_new / k)
    }

    # Step 3: k
    log_pk <- numeric(K_max)
    for (kk in 1:K_max) {
      jj <- pmin(pmax(ceiling(Z * kk), 1), kk)
      log_pk[kk] <- sum(dbeta(theta, jj, kk + 1 - jj, log = TRUE))
    }
    log_pk <- log_pk - max(log_pk)
    pk <- exp(log_pk); pk <- pk / sum(pk)
    k <- sample.int(K_max, 1, prob = pk)
    j_lab <- pmin(pmax(ceiling(Z * k), 1), k)

    # Step 4: (alpha, beta) via conjugate bivariate normal
    X_mat <- cbind(1, theta)
    prior_prec <- diag(c(1/alpha_prior_var, 1/beta_prior_var))
    prior_mean_vec <- c(alpha_prior_mean, beta_prior_mean)
    post_prec <- prior_prec + crossprod(X_mat) / sigma2
    post_var  <- solve(post_prec)
    post_mean <- post_var %*% (prior_prec %*% prior_mean_vec +
                               crossprod(X_mat, logN) / sigma2)
    ab_draw <- post_mean + t(chol(post_var)) %*% rnorm(2)
    alpha_op <- ab_draw[1]; beta_op <- ab_draw[2]

    # Step 5: sigma^2 via conjugate inverse-gamma
    resid <- logN - alpha_op - beta_op * theta
    post_a <- sigma2_prior_a + nn / 2
    post_b <- sigma2_prior_b + sum(resid^2) / 2
    sigma2 <- 1 / rgamma(1, post_a, post_b)

    # Step 6: M
    r <- length(unique(j_lab))
    log_pM <- sapply(M_grid, function(Mv) {
      r * log(Mv) - sum(log(Mv + 0:(nn - 1)))
    })
    log_pM <- log_pM - max(log_pM)
    pM <- exp(log_pM); pM <- pM / sum(pM)
    M <- M_grid[sample.int(length(M_grid), 1, prob = pM)]

    k_trace[iter]      <- k
    M_trace[iter]      <- M
    alpha_trace[iter]  <- alpha_op
    beta_trace[iter]   <- beta_op
    sigma2_trace[iter] <- sigma2

    if (iter > n_burn) {
      n_kept <- n_kept + 1
      nj <- tabulate(j_lab, nbins = k)
      W_rb <- (M / k + nj) / (M + nn)
      for (g in seq_along(theta_grid)) {
        density_sum[g] <- density_sum[g] +
          sum(W_rb * dbeta(theta_grid[g], 1:k, k + 1 - (1:k)))
      }
      theta_hat_sum <- theta_hat_sum + theta
    }

    if (iter %% 5000 == 0) {
      cat(sprintf("  iter %d/%d  k=%d  M=%.1f  alpha=%.2f  beta=%.2f  sigma2=%.2f",
                  iter, n_total, k, M, alpha_op, beta_op, sigma2))
      cat(sprintf("  theta_acc=%.1f%%\n", 100 * theta_accept / theta_total))
    }
  }

  cat(sprintf("  Final theta acceptance rate: %.1f%%\n",
              100 * theta_accept / theta_total))
  post_burn <- (n_burn + 1):n_total
  list(
    theta_grid     = theta_grid,
    density_est    = density_sum / n_kept,
    theta_hat      = theta_hat_sum / n_kept,
    k_samples      = k_trace[post_burn],
    M_samples      = M_trace[post_burn],
    alpha_samples  = alpha_trace[post_burn],
    beta_samples   = beta_trace[post_burn],
    sigma2_samples = sigma2_trace[post_burn]
  )
}
```

## 1.3 Comparison Estimators

```{r comparison-estimators}
# James-Stein on arcsine-square-root scale
stein_estimator <- function(y, N) {
  n <- length(y)
  x <- asin(sqrt((y + 0.25) / (N + 0.5)))
  x_bar <- mean(x)
  sig2 <- 1 / (4 * N)
  SS <- sum((x - x_bar)^2)
  B <- max(0, 1 - (n - 3) * mean(sig2) / SS)
  mu_s <- x_bar + B * (x - x_bar)
  sin(mu_s)^2
}
```

## 1.4 TSE* Computation (Brown 2008)

`TSE*` normalizes total squared error of an estimator by the naive estimator's
error, so `TSE* = 1` corresponds to the naive (MLE) and smaller values are better.

```{r tse-star}
compute_tse_star <- function(theta_hat, y1, N1, y2, N2) {
  val <- which(N2 >= 11)  # validation set: players with >= 11 second-half at-bats
  x1 <- asin(sqrt((y1[val] + 0.25) / (N1[val] + 0.5)))
  x2 <- asin(sqrt((y2[val] + 0.25) / (N2[val] + 0.5)))
  sig2 <- 1 / (4 * N2[val])
  mu_hat <- asin(sqrt(theta_hat[val]))
  tse_naive <- sum((x1 - x2)^2) - sum(sig2)
  tse_est   <- sum((mu_hat - x2)^2) - sum(sig2)
  tse_est / tse_naive
}
```


# Section 2: Reproduction of Zhang & Liu (2012), Table 1

We reproduce the four simulation scenarios in Zhang and Liu's Table 1.

```{r table1-simulations, cache=TRUE}
# Scenario 1: Beta(8,8), n=200, N ~ Uniform(100,200)
set.seed(100)
n1 <- 200
theta1 <- rbeta(n1, 8, 8)
N1_sim <- sample(100:200, n1, replace = TRUE)
y1_sim <- rbinom(n1, N1_sim, theta1)
cat("Scenario 1: Beta(8,8)\n")
res1 <- bdp_mcmc(y1_sim, N1_sim, n_iter = 15000, n_burn = 5000, K_max = 200)

# Scenario 2: Bimodal, n=200, N=100
set.seed(200)
n2 <- 200
comp2 <- rbinom(n2, 1, 0.5)
theta2 <- ifelse(comp2 == 1, rbeta(n2, 60, 10), rbeta(n2, 10, 60))
N2_sim <- rep(100L, n2)
y2_sim <- rbinom(n2, N2_sim, theta2)
cat("Scenario 2: Bimodal\n")
res2 <- bdp_mcmc(y2_sim, N2_sim, n_iter = 15000, n_burn = 5000, K_max = 200)

# Scenario 3: Exp(8) truncated at 1, n=200, N=100
set.seed(100)
n3 <- 200
u3 <- runif(n3)
theta3 <- -log(1 - u3 * (1 - exp(-8))) / 8
N3_sim <- rep(100L, n3)
y3_sim <- rbinom(n3, N3_sim, theta3)
cat("Scenario 3: Exp(8) truncated\n")
res3 <- bdp_mcmc(y3_sim, N3_sim, n_iter = 15000, n_burn = 5000,
                 K_max = 200, seed = 1100)

# Scenario 4: Mixture of Uniforms, n=100, N=100
set.seed(400)
n4 <- 100
comp4 <- rbinom(n4, 1, 0.5)
theta4 <- ifelse(comp4 == 1, runif(n4, 0.25, 0.50), runif(n4, 0.75, 1.00))
N4_sim <- rep(100L, n4)
y4_sim <- rbinom(n4, N4_sim, theta4)
cat("Scenario 4: Mix Uniforms\n")
res4 <- bdp_mcmc(y4_sim, N4_sim, n_iter = 15000, n_burn = 5000, K_max = 200)
```

```{r table1-print}
cat("========== TABLE 1 REPRODUCTION ==========\n")
cat(sprintf("%-30s %12s %12s\n", "Distribution", "Stein", "BDP"))
cat(paste(rep("-", 60), collapse = ""), "\n")

for (info in list(
  list("Beta(8,8)",     theta1, y1_sim, N1_sim, res1),
  list("Bimodal",       theta2, y2_sim, N2_sim, res2),
  list("Exp(8) trunc",  theta3, y3_sim, N3_sim, res3),
  list("Mix Uniforms",  theta4, y4_sim, N4_sim, res4)
)) {
  nm <- info[[1]]; th <- info[[2]]; yy <- info[[3]]
  NN <- info[[4]]; rr <- info[[5]]
  ase <- c(
    Stein = mean((stein_estimator(yy, NN) - th)^2),
    BDP   = mean((rr$theta_hat - th)^2)
  )
  cat(sprintf("%-30s %12.2e %12.2e\n", nm, ase[1], ase[2]))
}

cat("\nReference (Zhang & Liu 2012, Table 1):\n")
cat("Beta(8,8):    Stein=1.53e-3  BDP=1.53e-3\n")
cat("Bimodal:      Stein=1.17e-3  BDP=6.80e-4\n")
cat("Exp(8):       Stein=7.75e-4  BDP=7.20e-4\n")
cat("Mix Unif:     Stein=1.45e-3  BDP=1.04e-3\n")
```


# Section 3: Baseball Application

## 3.1 Load Brown's 2005 MLB Data

First half = April-June, second half = July-October. Filter: `N1 >= 11`.

```{r load-baseball-data}
raw <- read.csv("data/monthly_data_2005_commas.TXT",
                stringsAsFactors = FALSE)

raw$N1 <- raw$AB.4. + raw$AB.5. + raw$AB.6.
raw$H1 <- raw$H.4.  + raw$H.5.  + raw$H.6.
raw$N2 <- raw$AB.7. + raw$AB.8. + raw$AB.9.10.
raw$H2 <- raw$H.7.  + raw$H.8.  + raw$H.9.10.

brown_data <- raw[raw$N1 >= 11, ]
cat(sprintf("Brown data: %d players, %d pitchers, %d non-pitchers\n",
            nrow(brown_data), sum(brown_data$Pitcher. == 1),
            sum(brown_data$Pitcher. == 0)))

y_bb  <- brown_data$H1
N_bb  <- brown_data$N1
y2_bb <- brown_data$H2
N2_bb <- brown_data$N2
is_pit <- as.logical(brown_data$Pitcher.)
x1_bb <- asin(sqrt((y_bb + 0.25) / (N_bb + 0.5)))

cat(sprintf("N1 range: [%d, %d]\n", min(N_bb), max(N_bb)))
cat(sprintf("R^2(x_i, N_i): %.3f  (Paper: 0.247)\n", cor(x1_bb, N_bb)^2))
```

## 3.2 Standard BDP at K_max = 200

```{r run-standard-bdp-baseball, cache=TRUE}
res_bb <- bdp_mcmc(y_bb, N_bb, n_iter = 20000, n_burn = 5000, K_max = 200)
tse_std_bb <- compute_tse_star(res_bb$theta_hat, y_bb, N_bb, y2_bb, N2_bb)
tse_stein_bb <- compute_tse_star(stein_estimator(y_bb, N_bb),
                                  y_bb, N_bb, y2_bb, N2_bb)

cat(sprintf("Standard BDP TSE* = %.3f  (Paper: 0.663)\n", tse_std_bb))
cat(sprintf("James-Stein TSE*  = %.3f  (Paper: 0.525)\n", tse_stein_bb))
```

## 3.3 Informative BDP at K_max = 1200 (HEADLINE RESULT)

The K_max sensitivity sweep in Section 3.5 establishes that K_max = 1200 is the
smallest value at which the posterior of `k` is not materially constrained by
the boundary. We use this as the headline result.

```{r run-informative-bdp-baseball-K1200, cache=TRUE}
# Headline result: K_max = 1200
res_ln_K1200 <- informative_bdp_lognormal(y_bb, N_bb,
                                           n_iter = 20000, n_burn = 5000,
                                           K_max = 1200, seed = 42)
tse_inf_K1200 <- compute_tse_star(res_ln_K1200$theta_hat, y_bb, N_bb, y2_bb, N2_bb)
beta_mean_K1200 <- mean(res_ln_K1200$beta_samples)
beta_ci_K1200   <- quantile(res_ln_K1200$beta_samples, c(0.025, 0.975))

cat("\n========== BASEBALL HEADLINE RESULT ==========\n")
cat(sprintf("Informative BDP at K_max = 1200:\n"))
cat(sprintf("  TSE*       = %.3f\n", tse_inf_K1200))
cat(sprintf("  beta_hat   = %.2f\n", beta_mean_K1200))
cat(sprintf("  95%% CrI    = [%.2f, %.2f]\n", beta_ci_K1200[1], beta_ci_K1200[2]))
cat(sprintf("  alpha_hat  = %.2f\n", mean(res_ln_K1200$alpha_samples)))
cat(sprintf("  sigma2_hat = %.3f\n", mean(res_ln_K1200$sigma2_samples)))
```

## 3.4 Pitcher / Non-Pitcher Decomposition

```{r pitcher-decomposition}
# Validation pitchers: those with N2 >= 11 AND Pitcher. == 1
val_idx  <- which(N2_bb >= 11)
val_pit  <- val_idx[is_pit[val_idx]]
val_nonp <- val_idx[!is_pit[val_idx]]

cat(sprintf("Validation set: %d pitchers, %d non-pitchers\n",
            length(val_pit), length(val_nonp)))

# Aggregate squared error on raw probability scale
sq_err <- function(theta_hat, idx) {
  mu_hat <- asin(sqrt(theta_hat[idx]))
  x2     <- asin(sqrt((y2_bb[idx] + 0.25) / (N2_bb[idx] + 0.5)))
  sum((mu_hat - x2)^2)
}

sse_std_pit  <- sq_err(res_bb$theta_hat,       val_pit)
sse_inf_pit  <- sq_err(res_ln_K1200$theta_hat, val_pit)
sse_std_nonp <- sq_err(res_bb$theta_hat,       val_nonp)
sse_inf_nonp <- sq_err(res_ln_K1200$theta_hat, val_nonp)

cat("\n========== PITCHER DECOMPOSITION ==========\n")
cat(sprintf("Pitchers (n=%d):\n", length(val_pit)))
cat(sprintf("  Std BDP SSE = %.2f, mean theta_hat = %.3f\n",
            sse_std_pit, mean(res_bb$theta_hat[val_pit])))
cat(sprintf("  Inf BDP SSE = %.2f, mean theta_hat = %.3f\n",
            sse_inf_pit, mean(res_ln_K1200$theta_hat[val_pit])))
cat(sprintf("  Reduction:   %.1f%%\n", (sse_std_pit - sse_inf_pit) / sse_std_pit * 100))

cat(sprintf("\nNon-pitchers (n=%d):\n", length(val_nonp)))
cat(sprintf("  Std BDP SSE = %.2f, mean theta_hat = %.3f\n",
            sse_std_nonp, mean(res_bb$theta_hat[val_nonp])))
cat(sprintf("  Inf BDP SSE = %.2f, mean theta_hat = %.3f\n",
            sse_inf_nonp, mean(res_ln_K1200$theta_hat[val_nonp])))
```

## 3.5 K_max Sensitivity Sweep (Supplement D.1, Table D.1)

The posterior of `k` concentrates near the boundary when `K_max = 200`
(22.6% of mass at the boundary). We sweep `K_max` to confirm robustness.

NOTE: Each run takes 10-30 minutes. Skip with `eval=FALSE` and load saved
results if available.

```{r kmax-sweep-baseball, eval=FALSE}
# K_max = 200 (already computed above as res_ln_K200; rerun for clarity)
res_ln_K200  <- informative_bdp_lognormal(y_bb, N_bb, K_max = 200,
                                           n_iter = 20000, n_burn = 5000, seed = 42)
res_ln_K400  <- informative_bdp_lognormal(y_bb, N_bb, K_max = 400,
                                           n_iter = 20000, n_burn = 5000, seed = 42)
res_ln_K800  <- informative_bdp_lognormal(y_bb, N_bb, K_max = 800,
                                           n_iter = 20000, n_burn = 5000, seed = 42)
# K_max = 1200 already computed as res_ln_K1200

# Save all four for figures and table
save(res_bb, res_ln_K200, res_ln_K400, res_ln_K800, res_ln_K1200,
     file = "results/baseball_kmax_sweep.RData")
```

```{r kmax-sweep-table, eval=FALSE}
# Print Table D.1
make_kmax_row <- function(res, K) {
  tse <- compute_tse_star(res$theta_hat, y_bb, N_bb, y2_bb, N2_bb)
  bm  <- mean(res$beta_samples)
  bci <- quantile(res$beta_samples, c(0.025, 0.975))
  km  <- median(res$k_samples)
  bnd <- mean(res$k_samples == K) * 100
  c(K = K, TSE = tse, beta = bm, lo = bci[1], hi = bci[2],
    k_med = km, boundary_pct = bnd)
}

tab_kmax <- rbind(
  make_kmax_row(res_ln_K200,  200),
  make_kmax_row(res_ln_K400,  400),
  make_kmax_row(res_ln_K800,  800),
  make_kmax_row(res_ln_K1200, 1200)
)
print(tab_kmax)
```


# Section 4: Stack Overflow Application

The Stack Overflow data is extracted via SQL from the public Stack Exchange
Data Explorer (see `stackoverflow_query.sql`). Pre-computed `(y_so, N_so, y2_so, N2_so)`
are loaded from disk; the query and counts are documented in the manuscript
Supplement G.

```{r so-load, eval=FALSE}
# Load from pre-computed file; replace with your local path
load("data/so_data_2023.RData")
cat(sprintf("Stack Overflow: n = %d users\n", length(y_so)))
cat(sprintf("N range: [%d, %d], median = %d\n",
            min(N_so), max(N_so), median(N_so)))

x1_so <- asin(sqrt((y_so + 0.25) / (N_so + 0.5)))
cat(sprintf("cor(X_i, log N_i) = %.3f\n", cor(x1_so, log(N_so))))
```

## 4.1 Headline Result at K_max = 200

```{r so-headline, eval=FALSE}
res_so_K200 <- informative_bdp_lognormal(y_so, N_so,
                                          n_iter = 20000, n_burn = 5000,
                                          K_max = 200, seed = 123)
res_so_std  <- bdp_mcmc(y_so, N_so, n_iter = 20000, n_burn = 5000, K_max = 200)

tse_so_inf <- compute_tse_star(res_so_K200$theta_hat, y_so, N_so, y2_so, N2_so)
tse_so_std <- compute_tse_star(res_so_std$theta_hat,  y_so, N_so, y2_so, N2_so)
tse_so_js  <- compute_tse_star(stein_estimator(y_so, N_so),
                                y_so, N_so, y2_so, N2_so)

cat(sprintf("Naive MLE        TSE* = %.3f\n", 1.113))
cat(sprintf("James-Stein      TSE* = %.3f\n", tse_so_js))
cat(sprintf("Standard BDP     TSE* = %.3f\n", tse_so_std))
cat(sprintf("Informative BDP  TSE* = %.3f  (HEADLINE)\n", tse_so_inf))
cat(sprintf("  beta_hat = %.2f, 95%% CrI = [%.2f, %.2f]\n",
            mean(res_so_K200$beta_samples),
            quantile(res_so_K200$beta_samples, 0.025),
            quantile(res_so_K200$beta_samples, 0.975)))
```

## 4.2 K_max Sensitivity Check on Stack Overflow

The Stack Overflow posterior of `k` places only 3.1% mass at `K_max = 200`,
substantially lower than baseball's 22.6%. We rerun at `K_max = 400` to
confirm stability.

```{r so-kmax-check, eval=FALSE}
res_so_K400 <- informative_bdp_lognormal(y_so, N_so,
                                          n_iter = 20000, n_burn = 5000,
                                          K_max = 400, seed = 123)
tse_so_K400  <- compute_tse_star(res_so_K400$theta_hat, y_so, N_so, y2_so, N2_so)
beta_so_K400 <- mean(res_so_K400$beta_samples)
bci_so_K400  <- quantile(res_so_K400$beta_samples, c(0.025, 0.975))
bnd_so_K400  <- mean(res_so_K400$k_samples == 400) * 100

cat(sprintf("K=200: TSE*=%.4f, beta=%.2f, boundary=%.1f%%\n",
            tse_so_inf, mean(res_so_K200$beta_samples),
            mean(res_so_K200$k_samples == 200) * 100))
cat(sprintf("K=400: TSE*=%.4f, beta=%.2f, boundary=%.1f%%\n",
            tse_so_K400, beta_so_K400, bnd_so_K400))
cat("Conclusion: results stable; K=200 retained as headline.\n")
```


# Section 5: Simulation Study (5 Scenarios, 30 Replicates)

This is the simulation study underlying the manuscript's Table 1 (main text).
WARNING: full run takes ~6-10 hours. Skip with `eval=FALSE` and load saved
results if available.

```{r sim-study-setup}
gen_beta88  <- function(n) rbeta(n, 8, 8)
gen_bimodal <- function(n) {
  comp <- rbinom(n, 1, 0.5)
  ifelse(comp == 1, rbeta(n, 60, 10), rbeta(n, 10, 60))
}

scenarios <- list(
  S1 = list(name = "S1: Beta(8,8), beta=0",
            F_gen = gen_beta88,  beta_true = 0, n = 200, alpha_true = 4.0),
  S2 = list(name = "S2: Beta(8,8), beta=2",
            F_gen = gen_beta88,  beta_true = 2, n = 200, alpha_true = 4.0),
  S3 = list(name = "S3: Bimodal, beta=0",
            F_gen = gen_bimodal, beta_true = 0, n = 200, alpha_true = 4.0),
  S4 = list(name = "S4: Bimodal, beta=2",
            F_gen = gen_bimodal, beta_true = 2, n = 200, alpha_true = 4.0),
  S5 = list(name = "S5: Bimodal, beta=3, n=100",
            F_gen = gen_bimodal, beta_true = 3, n = 100, alpha_true = 4.0)
)

N_REP     <- 30
SIM_NITER <- 8000
SIM_NBURN <- 2000
SIM_KMAX  <- 80
```

```{r sim-study-run, eval=FALSE}
run_one_rep <- function(sc, rep_seed) {
  set.seed(rep_seed)
  n <- sc$n
  theta_true <- sc$F_gen(n)
  N_obs <- rpois(n, exp(sc$alpha_true + sc$beta_true * theta_true))
  N_obs <- pmax(N_obs, 1L)
  y_obs <- rbinom(n, N_obs, theta_true)

  res_std <- bdp_mcmc(y_obs, N_obs, n_iter = SIM_NITER, n_burn = SIM_NBURN,
                      K_max = SIM_KMAX, seed = rep_seed + 1000)
  res_inf <- informative_bdp_lognormal(y_obs, N_obs,
                                        n_iter = SIM_NITER, n_burn = SIM_NBURN,
                                        K_max = SIM_KMAX, seed = rep_seed + 2000)

  c(BDP_std  = mean((res_std$theta_hat - theta_true)^2),
    BDP_inf  = mean((res_inf$theta_hat - theta_true)^2),
    beta_hat = mean(res_inf$beta_samples))
}

all_results <- list()
for (sc_name in names(scenarios)) {
  sc <- scenarios[[sc_name]]
  cat(sprintf("\n========== %s ==========\n", sc$name))
  results_mat <- matrix(NA, nrow = N_REP, ncol = 3)
  colnames(results_mat) <- c("BDP_std", "BDP_inf", "beta_hat")
  for (rep in 1:N_REP) {
    cat(sprintf("  Rep %d/%d ... ", rep, N_REP))
    t0 <- Sys.time()
    tryCatch({
      results_mat[rep, ] <- run_one_rep(sc,
        rep_seed = rep * 100 + as.integer(sc$beta_true * 10))
      cat(sprintf("done (%.0fs)\n",
                  as.numeric(difftime(Sys.time(), t0, units = "secs"))))
    }, error = function(e) cat(sprintf("ERROR: %s\n", e$message)))
  }
  all_results[[sc_name]] <- results_mat
}
save(all_results, scenarios, file = "results/sim_30rep_results.RData")
```

```{r sim-display-results}
if (file.exists("results/sim_30rep_results.RData")) {
  load("results/sim_30rep_results.RData")

  cat("================================================================\n")
  cat("  30-REPLICATE SIMULATION RESULTS (mean ASE; sd in parens)\n")
  cat("================================================================\n")
  cat(sprintf("%-30s %16s %16s %10s %10s\n",
              "Scenario", "Std BDP", "Inf BDP", "Delta(%)", "beta_hat"))
  cat(paste(rep("-", 90), collapse = ""), "\n")

  for (sc_name in names(scenarios)) {
    sc <- scenarios[[sc_name]]
    rm <- all_results[[sc_name]]
    valid <- !is.na(rm[, 1])
    m <- colMeans(rm[valid, , drop = FALSE])
    s <- apply(rm[valid, , drop = FALSE], 2, sd)
    pct <- (m["BDP_std"] - m["BDP_inf"]) / m["BDP_std"] * 100
    cat(sprintf("%-30s  %.2e (%.1e)  %.2e (%.1e)  %+8.1f  %8.2f\n",
                sc$name,
                m["BDP_std"], s["BDP_std"],
                m["BDP_inf"], s["BDP_inf"],
                pct, m["beta_hat"]))
  }
} else {
  cat("results/sim_30rep_results.RData not found.\n")
  cat("Run the simulation study chunk above to generate it.\n")
}
```


# Section 6: Structural Diagnostic (Supplement E, Tests A-D)

Targeted experiments to diagnose the source of the ~18% cost in scenario S3
(bimodal, beta=0). The unifying finding: the cost depends only on the magnitude
of beta, with a crossover near `|beta| = 1`.

```{r diagnostic-setup}
gen_mild_bimodal <- function(n) {
  comp <- rbinom(n, 1, 0.5)
  ifelse(comp == 1, rbeta(n, 20, 10), rbeta(n, 10, 20))
}

run_comparison <- function(F_gen, label, n = 200, beta = 0,
                           n_iter = 8000, n_burn = 2000, K_max = 80, seed = 42) {
  set.seed(seed)
  theta_true <- F_gen(n)
  N_obs <- rpois(n, exp(4.0 + beta * theta_true))
  N_obs <- pmax(N_obs, 1)
  y_obs <- rbinom(n, N_obs, theta_true)

  res_std <- bdp_mcmc(y_obs, N_obs, n_iter = n_iter, n_burn = n_burn,
                      K_max = K_max, seed = seed)
  res_inf <- informative_bdp_lognormal(y_obs, N_obs, n_iter = n_iter,
                                        n_burn = n_burn, K_max = K_max, seed = seed)

  ase_std <- mean((res_std$theta_hat - theta_true)^2)
  ase_inf <- mean((res_inf$theta_hat - theta_true)^2)
  pct <- (ase_std - ase_inf) / ase_std * 100
  beta_hat <- mean(res_inf$beta_samples)

  cat(sprintf("  %-40s std=%.4e inf=%.4e gap=%+5.1f%% beta=%.2f\n",
              label, ase_std, ase_inf, pct, beta_hat))
  c(ase_std = ase_std, ase_inf = ase_inf, pct = pct, beta_hat = beta_hat)
}
```

```{r diagnostic-test-A, eval=FALSE}
# Test A: Distribution shape (all beta = 0)
cat("\n=== TEST A: Distribution shape, beta=0 ===\n")
test_A_results <- list(
  unimodal      = run_comparison(gen_beta88,        "Beta(8,8) unimodal", beta = 0),
  mild_bimodal  = run_comparison(gen_mild_bimodal,  "Mild bimodal",       beta = 0),
  uniform       = run_comparison(function(n) runif(n), "Uniform(0,1)",    beta = 0),
  strong_bimod  = run_comparison(gen_bimodal,       "Strong bimodal",     beta = 0)
)
```

```{r diagnostic-test-B, eval=FALSE}
# Test B: Crossover beta (strong bimodal)
cat("\n=== TEST B: Crossover beta (strong bimodal) ===\n")
test_B_results <- list()
for (b in c(0, 0.25, 0.5, 0.75, 1.0, 1.5, 2.0)) {
  key <- sprintf("beta_%.2f", b)
  test_B_results[[key]] <- run_comparison(gen_bimodal,
                                            sprintf("beta = %.2f", b),
                                            beta = b)
}
```

```{r diagnostic-test-C, eval=FALSE}
# Test C: Sampler length (strong bimodal, beta = 0)
cat("\n=== TEST C: Sampler length (bimodal, beta=0) ===\n")
test_C_results <- list()
for (ni in c(5000, 10000, 20000, 40000)) {
  key <- sprintf("iter_%d", ni)
  test_C_results[[key]] <- run_comparison(gen_bimodal,
                                            sprintf("n_iter = %d", ni),
                                            n_iter = ni, n_burn = ni %/% 4)
}
```

```{r diagnostic-test-D, eval=FALSE}
# Test D: Sample size (strong bimodal, beta = 0)
cat("\n=== TEST D: Sample size (bimodal, beta=0) ===\n")
test_D_results <- list()
for (nn in c(100, 200, 500)) {
  key <- sprintf("n_%d", nn)
  test_D_results[[key]] <- run_comparison(gen_bimodal,
                                            sprintf("n = %d", nn),
                                            n = nn)
}
```


# Section 7: Errors-in-Variables Verification (Supplement E.2)

Numerical check of the attenuation correction on Brown's baseball data.

```{r eiv-baseball}
# Empirical attenuation factor
mle_bb <- y_bb / N_bb
N_clean <- N_bb[N_bb > 0]
mle_clean <- y_bb[N_bb > 0] / N_clean
var_mle <- var(mle_clean)
e_var_mle_given_theta <- mean(mle_clean * (1 - mle_clean) / N_clean)
var_theta <- var_mle - e_var_mle_given_theta
kappa <- var_theta / (var_theta + e_var_mle_given_theta)

# OLS regression of log(N) on the noisy MLE
fit_naive <- lm(log(N_bb) ~ mle_bb)
beta_ols <- coef(fit_naive)[2]

# Attenuation-corrected estimate
beta_corrected <- beta_ols / kappa

cat("========== ERRORS-IN-VARIABLES (Baseball) ==========\n")
cat(sprintf("Var(MLE)                     = %.3e\n", var_mle))
cat(sprintf("E[theta(1-theta)/N]          = %.3e\n", e_var_mle_given_theta))
cat(sprintf("Implied Var(theta)           = %.3e\n", var_theta))
cat(sprintf("Attenuation kappa            = %.3f\n", kappa))
cat(sprintf("OLS beta (noisy regressor)   = %.2f\n", beta_ols))
cat(sprintf("Attenuation-corrected beta   = %.2f\n", beta_corrected))
cat(sprintf("Posterior mean beta (MCMC)   = %.2f\n", beta_mean_K1200))

# Independent verification: OLS on noise-free posterior means
fit_clean <- lm(log(N_bb) ~ res_ln_K1200$theta_hat)
cat(sprintf("OLS on posterior means       = %.2f\n", coef(fit_clean)[2]))
```


# Section 8: Figures for the Manuscript

All figures are written to `figs/` for inclusion in the LaTeX manuscript.

```{r setup-figs-dir}
if (!dir.exists("figs")) dir.create("figs")
```

## 8.1 Figure D.1: Trace Plots (Supplement D)

```{r fig-D1-traces, eval=TRUE}
pdf("figs/FigD1_traces.pdf", width = 7, height = 5)
par(mfrow = c(2, 2), family = "serif",
    mar = c(4, 4, 2.5, 1), mgp = c(2.5, 0.8, 0), tcl = -0.3)

iter_idx <- seq_along(res_ln_K1200$alpha_samples)

plot(iter_idx, res_ln_K1200$alpha_samples, type = "l",
     col = "steelblue", lwd = 0.5,
     xlab = "Iteration (post burn-in)", ylab = expression(alpha),
     main = expression("(a) Trace of " * alpha))
abline(h = mean(res_ln_K1200$alpha_samples), col = "red", lwd = 1.2)

plot(iter_idx, res_ln_K1200$beta_samples, type = "l",
     col = "darkgreen", lwd = 0.5,
     xlab = "Iteration (post burn-in)", ylab = expression(beta),
     main = expression("(b) Trace of " * beta))
abline(h = mean(res_ln_K1200$beta_samples), col = "red", lwd = 1.2)

plot(iter_idx, res_ln_K1200$sigma2_samples, type = "l",
     col = "purple", lwd = 0.5,
     xlab = "Iteration (post burn-in)", ylab = expression(sigma^2),
     main = expression("(c) Trace of " * sigma^2))
abline(h = mean(res_ln_K1200$sigma2_samples), col = "red", lwd = 1.2)

plot(iter_idx, res_ln_K1200$k_samples, type = "l",
     col = "darkorange", lwd = 0.5,
     xlab = "Iteration (post burn-in)", ylab = "k",
     main = "(d) Trace of k")
abline(h = mean(res_ln_K1200$k_samples), col = "red", lwd = 1.2)

dev.off()
cat("Saved: figs/FigD1_traces.pdf\n")
```

## 8.2 Figure E.1: Crossover Plot (Supplement E, Test B)

```{r fig-E1-crossover}
# These ASE values come from your Test B run (Section 6)
beta_grid <- c(0, 0.25, 0.5, 0.75, 1.0, 1.5, 2.0)
ase_std   <- c(1.1783e-3, 1.0734e-3, 1.0496e-3, 9.8628e-4,
               8.9387e-4, 7.8997e-4, 6.5734e-4)
ase_inf   <- c(1.4506e-3, 1.2855e-3, 1.2030e-3, 1.0936e-3,
               9.1832e-4, 7.4697e-4, 5.1091e-4)
gap_pct   <- (ase_std - ase_inf) / ase_std * 100

pdf("figs/FigE1_crossover.pdf", width = 5.5, height = 4)
par(family = "serif", mar = c(4, 4, 2.5, 1),
    mgp = c(2.5, 0.8, 0), tcl = -0.3)

plot(beta_grid, gap_pct, type = "b",
     pch = 16, cex = 1.3, lwd = 1.5, col = "navy",
     xlab = expression("True " * beta),
     ylab = "ASE gap (%), Informative BDP vs Standard BDP",
     main = expression("Crossover under bimodal F"),
     ylim = c(-30, 30))
abline(h = 0, col = "gray50", lwd = 1)

text(1.0, -27, expression("Crossover at " * beta %~~% 1),
     cex = 0.9, col = "gray30")
arrows(1.0, -24, 1.0, -3, length = 0.08, col = "gray30")
text(0.4, 20, "BDP (std) better",       cex = 0.9, col = "darkred")
text(1.6, 20, "Informative BDP better", cex = 0.9, col = "darkgreen")

dev.off()
cat("Saved: figs/FigE1_crossover.pdf\n")
```

## 8.3 Figure E.2: Shrinkage Visualization (Supplement E)

The mechanism figure: how Standard BDP over-shrinks pitchers, while
Informative BDP correctly retains them near low MLE values.

```{r fig-E2-shrinkage}
mle_bb_cc <- (y_bb + 0.5) / (N_bb + 1)  # MLE with continuity correction

pdf("figs/FigE2_shrinkage.pdf", width = 11, height = 5.5)
par(mfrow = c(1, 2),
    mar = c(4.5, 4.5, 2.5, 1),
    cex.lab = 1.15, cex.main = 1.2, cex.axis = 1.05)

# Panel (a): Standard BDP
plot(mle_bb_cc, res_bb$theta_hat,
     pch = ifelse(is_pit, 17, 1),
     col = ifelse(is_pit, "firebrick", "steelblue"),
     cex = ifelse(is_pit, 1.1, 0.75),
     xlab = expression(paste("First-half MLE  ", y[i]/N[i])),
     ylab = expression(paste("Posterior mean  ", hat(theta)[i])),
     main = "(a) Standard BDP",
     xlim = c(0, 0.45), ylim = c(0.05, 0.40))
abline(0, 1, lty = 2, col = "gray50", lwd = 1)
abline(h = mean(res_bb$theta_hat[is_pit]),  col = "firebrick", lty = 3, lwd = 1)
abline(h = mean(res_bb$theta_hat[!is_pit]), col = "steelblue", lty = 3, lwd = 1)
legend("topleft",
       legend = c("Pitchers", "Non-pitchers", "MLE (no shrinkage)"),
       pch = c(17, 1, NA), lty = c(NA, NA, 2),
       col = c("firebrick", "steelblue", "gray50"),
       bty = "n", cex = 0.95)

# Panel (b): Informative BDP at K_max = 1200
plot(mle_bb_cc, res_ln_K1200$theta_hat,
     pch = ifelse(is_pit, 17, 1),
     col = ifelse(is_pit, "firebrick", "steelblue"),
     cex = ifelse(is_pit, 1.1, 0.75),
     xlab = expression(paste("First-half MLE  ", y[i]/N[i])),
     ylab = expression(paste("Posterior mean  ", hat(theta)[i])),
     main = expression(paste("(b) Informative BDP, ", K[max], "=1200")),
     xlim = c(0, 0.45), ylim = c(0.05, 0.40))
abline(0, 1, lty = 2, col = "gray50", lwd = 1)
abline(h = mean(res_ln_K1200$theta_hat[is_pit]),  col = "firebrick", lty = 3, lwd = 1)
abline(h = mean(res_ln_K1200$theta_hat[!is_pit]), col = "steelblue", lty = 3, lwd = 1)

dev.off()
cat("Saved: figs/FigE2_shrinkage.pdf\n")
```

## 8.4 Figure F.1: Density Reproduction of Zhang & Liu Figure 2

```{r fig-F1-density}
pdf("figs/FigF1_density_reproduction.pdf", width = 7, height = 6)
par(mfrow = c(2, 2), family = "serif",
    mar = c(4, 4, 2.5, 1), mgp = c(2.5, 0.8, 0), tcl = -0.3)

true_densities <- list(
  function(x) dbeta(x, 8, 8),
  function(x) 0.5 * dbeta(x, 60, 10) + 0.5 * dbeta(x, 10, 60),
  function(x) 8 * exp(-8 * x) / (1 - exp(-8)) * (x >= 0 & x <= 1),
  function(x) ifelse(x >= 0.25 & x <= 0.5, 2, 0) +
              ifelse(x >= 0.75 & x <= 1.0, 2, 0)
)
names_sim <- c("(a) Beta(8,8)",
               "(b) 0.5 Beta(60,10) + 0.5 Beta(10,60)",
               "(c) Exp(8) truncated",
               "(d) 0.5 Unif(0.25,0.5) + 0.5 Unif(0.75,1)")
all_res <- list(res1, res2, res3, res4)

for (i in 1:4) {
  rr <- all_res[[i]]
  tg <- rr$theta_grid
  td <- true_densities[[i]](tg)
  yl <- c(0, max(c(rr$density_est, td), na.rm = TRUE) * 1.15)

  plot(tg, rr$density_est, type = "l", lwd = 2, col = "blue",
       main = names_sim[i],
       xlab = expression(theta), ylab = "Density",
       ylim = yl)
  lines(tg, td, col = "red", lwd = 2, lty = 2)
  legend("topright",
         c("BDP estimate", "Truth"),
         col = c("blue", "red"),
         lwd = 2, lty = c(1, 2),
         cex = 0.75, bty = "n")
}

dev.off()
cat("Saved: figs/FigF1_density_reproduction.pdf\n")
```
